---
title: "Assignment 11 - Respondent Bias"
author: "Statenfreude"
format:
  revealjs: 
    incremental: true
echo: false
---

```{r initial setup}
library(tidyverse); library(janitor); library(kableExtra)
theme_set(theme_minimal())
df <- read_csv("survey-data.csv") |> clean_names()

# gonna keep respondent just in case there are interesting questions about
# respondent behaviours.
df_long <- df |> select(respondent, starts_with("question")) |> 
  pivot_longer(cols=starts_with("question"), names_to="question", 
                            values_to="response")
```   

## Research Question
- What study strategies do university students think are effective and used frequently?  
- Verbal frequency and Likert scales were used to gauge which strategies used. 
- Many popular learning methods, highlighting & rereading are ineffective.  
- recall is a highly effective learning strategy but requires more effort

**Make it Stick**, $\text{Brown, Rudiger & McDonald,2014}$

## Acquiesence Bias
- Subjects respond more positively if there is an indication of what is a positive response
is included in the prompt.  
- We attempted to bias the question about recall.  
- The two questions representing passive learning methods were left as unbiased as possible.  

## Hypotheses
- Passive, less effective methods will be more frequent $\to$ Highlighting, rereading  
- Active, more effective method will be less frequent $\to$ Recall  
- Highlighting may be less frequent because of digitization  
- After bias, recall will be more frequent

## Results
- All distributions skewed  
- Highlighter distribution skewed positive, others negative
```{r histograms of question responses}
#| fig-width: 6
#| fig-height: 4
#| fig-align: "center"
question_names <- c(question_1 = "Highlighting",
                    question_2 = "Rereading",
                    question_3 = "Recall")

df_long |> ggplot(aes(x=response)) + 
  geom_bar(fill="steelblue") + facet_wrap(~question, labeller = as_labeller(question_names)) +
  labs(
    y="Frequency", x="Response",
    title="Learning method use among Univeristy of Wyoming students (n=40)"
  ) +
  theme(
    strip.text = element_text(size = 13, face="bold")
  )
```

## Results (continued)
- Large IQR for highlighting  
- Rereading most favoured learning method
- Recall notably preferred to highlighting

```{r results descriptive stats}
sum_stats <- rbind(
  summary(df$question_1),
  summary(df$question_2),
  summary(df$question_3)  
) 
rownames(sum_stats) <- c("Highlighting", "Rereading", "Recall")
sum_stats |> kable()
```

## Conclusion
- Recall (med=4) and rereading (med=5) very similar  
- Suggests bias pushed recall results higher  
- Highlighting (med=2) lowest $\to$ digitization  
- More difficult, active method performs better/the same after bias

# Thank you!


