---
title: "Assignment 11"
author: "Statenfreude"
format: html
execute: 
  warning: false
  messages: false
---

```{r initial setup}
library(tidyverse); library(janitor)
theme_set(theme_minimal())
df <- read_csv("survey-data.csv") |> clean_names()

# gonna keep respondent just in case there are interesting questions about
# respondent behaviours.
df_long <- df |> select(respondent, starts_with("question")) |> 
  pivot_longer(cols=starts_with("question"), names_to="question", 
               values_to="response")
```



## Questions

The form of respondent bias your group hopes to illustrate


## Expected results
We expected that the reported usage of passive methods of highlighting and
re-reading would be more frequent than recall, which is generally more taxing.
There was some concern that highlighting may be less relevant now that many
students only use electronic copies of texts and use fewer printed learning
materials.

## Results


a. the appropriate descriptive results to appropriately summarize each question separately
b. the results needed to evaluate your expected results
c. If not included in b, then "appropriately" correlate the results of your questions


Re-reading appears to be most widely used. Highlighting is not a common practice
among respondents.




```{r histograms of question responses}
#| fig-cap: "A comparison of learning method frequency of use among a sample of $n=40$ Univeristy of Wyoming students."
question_names <- c(question_1 = "Highlighting",
                    question_2 = "Re-reading",
                    question_3 = "Recall")

df_long |> ggplot(aes(x=response)) + 
  geom_bar() + facet_wrap(~question, labeller = as_labeller(question_names)) +
  labs(
    y="Frequency", x="Response"
  )

```


## Conclusions


